---
title: "prior"
author: "Rich"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: "css/my-css.css"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file samples the prior distribution to generate predictions. The main aim
is to check to see if the priors generate values that look sensible, given our
domain knowledge of the type of data.

From Gelman & co., page 10:

>Prior predictive checks are a useful tool to understand the implications of a 
prior distribution in the context of a generative model (Box, 1980, Gabry et al., 2019; 
see also Section 7.3 for details on how to work with prior distributions). 
In particular, because prior predictive checks make use of simulations from the 
model rather than observed data, they provide a way to refine the model without 
using the data multiple times.

And page 11 on generative models:

>Fully Bayesian data analysis requires a generative model—that is, a joint 
probability distribution for all the data and parameters. The point is subtle: 
Bayesian inference does not actually require the generative model; all it needs 
from the data is the likelihood, and different generative models can have the 
same likelihood. But Bayesian data analysis requires the generative model to be 
able to perform predictive simulation and model checking (Sections 2.4, 4.1, 
4.2, 6.1, and 6.2), and Bayesian workflow will consider a series of generative models.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
# install.packages("remotes")
# remotes::install_github("stan-dev/cmdstanr")
# 
# install.packages("devtools")
# devtools::install_github("jmgirard/standist")
# 
# install.packages(c("tidyverse", "RColorBrewer", "patchwork", "brms",
#                    "tidybayes", "bayesplot", "patchwork", "future", "faux"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel", "faux")

lapply(pkg, library, character.only = TRUE)
```

## settings ##

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```


The overall Bayesian workflow will be split up into several sections, as follows:

1. prior predictions.
2. planning for power/precision via simulation.
3. build a series of models and perform model checking and model comparison.
4. evaluate the posterior.

This script is part (1) above. 

## prior predictions ##

## section 1 - a simple dataset ##

- say something about prior preds. quote Gelman et al.

- describe the example design/s, which vary in complexity. dv=rt.

d1 - congr vs incon, no varying effects.

d2 - congr vs incon, varying intercepts and slopes for pid and items.

having two designs means that folks might follow along a little easier without
getting lost in varying effects (if they are not familiar with them).

Also make a note about the choice of distribution. e.g., RT data tends to be 
left skewed and gaussian models are a sub-optimal fit. Various other dists are better
e.g., link here: https://lindeloev.shinyapps.io/shiny-rt/. Shifted lognormal, 
for example, is one that we have used before in our own work. For simplicity and 
convenience, we will generate gaussian data here and use a gaussian model. 

simulate some data

```{r}
# define sample size
n = 100

# define the within pt conditions
within <- list(condition = c("congr", "incon"))

# define the means per condition
mu <- data.frame(
  condition = c(700, 750),
  row.names = within$condition
)

# add factor labels for plotting
vardesc <- c(condition = "condition")

d1 <- sim_design(within, 
                 n = n, mu = mu, sd = 25, r = .5,
                 empirical = TRUE, vardesc = vardesc, plot = TRUE)
head(d1)
```

make it into a longer format

```{r}
d1 <- d1 %>% 
  pivot_longer(-id,
               names_to = "condition",
               values_to = "rt")
head(d1)

# save initial data
write_csv(d1, "data/d1.csv")
```

sample the prior

formula

```{r}
formula = bf(rt ~ 1 + condition) 
```

check priors

```{r}
get_prior(data = d1,
          family = gaussian,
          formula)
```

visualise prior settings

in original units 

for the intercept

```{r}
visualize("normal(700, 200)", "normal(700, 150)", "normal(700, 100)", 
          "normal(700, 50)", 
          xlim = c(0, 1200))
```

ok, so based on what we know about these kinds of tasks 700,100 or 700,150 seems
pretty good for a weakly informative approach

for the effect of condition

```{r}
visualize("normal(0, 10)", "normal(0, 25)", "normal(0, 50)",
          xlim = c(-100, 100))
```

ok, based on this, 0,50 would do the job because effects over 100 or below -100
seem unlikely, given what we know about these tasks. Narrower could be better 
still tbh.


set priors

(you could also leave them unspecified and go with default priors, but I think
it's good to keep practicing the logic and though process behind them, so I try 
to set them explicitly whenever I can.)

```{r}
priors <- c(
  set_prior("normal(700, 150)", class = "Intercept"),
  set_prior("normal(0, 50)", class = "b"),
  set_prior("normal(0, 50)", class = "sigma")
)
```

build the model to sample the prior

```{r}
bp1 <- brm(data = d1,
      family = gaussian,
      formula = formula,
      prior = priors,
      seed = 1,
      sample_prior = "only")
summary(bp1)
```

1 divergent transition (out of 4000). But because we are using 200 observations
as a test, let's just ignore this for now.

let's take a look

```{r}
plot(bp1)
```

now let's take a look at the prior predictions

pp_check

```{r}
prior_dist <- pp_check(bp1, ndraws=100) +
  ggtitle("prior_predictions")
prior_dist
```


## section 2 - a more complicated and realistic dataset ##

simulate some data

```{r}
# define parameters
subj_n = 25  # number of subjects
item_n = 10  # number of items (10 faces x 8 repeats (half gaze left))
rep_n = 8 # number of trial repeats per item e.g., face1 is shown X times per pid 50% L, 50% C
b0 = 725      # intercept
b1 = 50      # fixed effect of condition
u0s_sd = 50   # random intercept SD for subjects
u1s_sd = 10   # random b1 slope SD for subjects
u0i_sd = 25   # random intercept SD for items 
u1i_sd = 5   # random b1 slope SD for items
r01s = 0.3   # correlation between random effects 0 and 1 for subjects
r01i = 0.3   # correlation between random effects 0 and 1 for items
sigma_sd = 50 # error SD

# set up data structure
d2 <- add_random(subj = subj_n, item = item_n, rep = rep_n) %>%
  # add and recode categorical variables
  add_within("item", gaze = c("left", "right")) %>%
  add_within("subj", condition = c("congr", "incong")) %>%
  add_contrast("condition", "anova", add_cols = TRUE, colnames = "cond") %>%
  # add random effects 
  add_ranef("subj", u0s = u0s_sd, u1s = u1s_sd, .cors = r01s) %>%
  add_ranef("item", u0i = u0i_sd, u1i = u1i_sd, .cors = r01i) %>%
  add_ranef(sigma = sigma_sd) %>%
  # calculate DV
  mutate(rt = b0 + u0s + u0i + (b1 + u1s + u1i) * cond + sigma)

head(d2)
str(d2)
summary(d2)

# save initial data
write_csv(d2, "data/d2.csv")
```

density plot

```{r}
ggplot(d2, aes(x=rt, fill=condition)) +
   geom_density(alpha = 0.3, colour = "darkgrey") +
   scale_fill_brewer(palette = "Dark2")+
   theme_bw()+
   theme(panel.grid = element_blank()) +
   theme(legend.position = "none") +
   ggtitle("rt by condition")
# ggsave ("figures/density.jpeg")
```

sample the prior

formula

```{r}
formula = bf(rt ~ 1 + cond + 
               (1 + cond | subj) +
               (1 + cond | item)) 
```

check priors

```{r}
get_prior(data = d2,
          family = gaussian,
          formula)
```

visualise prior settings

we did this above.

set priors

(you could also leave them unspecified and go with default priors, but I think
it's good to keep practicing the logic and though process behind them, so I try 
to set them explicitly whenever I can.)

```{r}
priors <- c(
  set_prior("normal(700, 150)", class = "Intercept"),
  set_prior("normal(0, 50)", class = "b"),
  set_prior("normal(0, 50)", class = "sigma"),
  set_prior("normal(0, 50)", class = "sd"),
  set_prior("lkj(2)", class = "cor")
)
```

build the model to sample the prior

```{r}
bp2 <- brm(data = d2,
      family = gaussian,
      formula = formula,
      prior = priors,
      seed = 1,
      sample_prior = "only")
summary(bp2)
```

that built with no problems.

let's take a look

```{r}
plot(bp2)
```

now let's take a look at the prior predictions

pp_check

```{r}
prior_dist2 <- pp_check(bp2, ndraws=100, type = "dens_overlay_grouped",
                        group = "cond") +
  ggtitle("prior_predictions")
prior_dist2
```